R Notebook

### 2021-07-11 draft implementation 2.Rmd
da finire**

>>>>

\
\textbf{Implementation}
\
Consider the linear Gaussian example of section XX, but this time with unknown variances $\tau^{2}$ and $\sigma^{2}$. Thus, let $\psi=(\sigma^{2},\tau^{2})$ be the unknown parameter vector and assign a gamma prior for its components,
\begin{align*}
\sigma^{2}  & \sim G(\alpha_{v},\beta_{v}) \\
\tau^{2}  & \sim G(\alpha_{w},\beta_{w})
\end{align*}
Alternatively, assign them a uniform prior if we have no knowledge of the hyperparameters. The algorithm follows the following steps.
\begin{algorithm} LWF for Random Walk plus Noise Model
\begin{itemize}
\item Initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$, $({\sigma^{2}}^{(1)},...,{\sigma^{2}}^{(N)})$ from $G(\alpha_{v},\beta_{v})$ and $({\tau^{2}}^{(1)},...,{\tau^{2}}^{(N)})$ from $G(\alpha_{w},\beta_{w})$. Set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$. Therefore $\psi^{(i)}=({\sigma^{2}}^{(i)},{\tau^{2}}^{(i)})$, and
$p(x_{0}|y_{0})=\sum_{i=1}^{N}w_{0}^{(i)}\delta_{(x_{0}^{(i)},\psi^{(i)})}$
\item For $t=1,...,n$:
\begin{enumerate}
\item Compute $$
\end{enumerate}

and $(\psi^{(1)},...,\psi^{(N)})$ from

Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item For $k=1,...,N$:
\begin{enumerate}
\item Draw $I_{k}$ with $P(I_{k}) \propto w_{t-1}^{(i)}f(y_{t}|g(x_{t-1}^{(i)})) $
\item Draw $x_{t}^{(k)} \sim N(x_{t-1}^{(I_{k})},\tau^2)$
\item Set  $\tilde{w}_{t}^{(k)} = \frac{f_{N}(y_{t}|x_{t}^{(k)})}{f_{N}(y_{t}|g(x_{t-1}^{(I_{k})}))}$
\end{enumerate}
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}

After having drown the hyperparameters  indipendently form their priors and having set $w_{0}^{(i)}=N^{-1}$, $i=1,...,N$, and $\hat{\pi}_{0}=\sum_{i=1}^{N}w_{0}^{(i)}\delta_{(x_{0}^{(i)},\psi^{(i)})}$, for t=1,..T 
\begin{itemize}
\item Compute $\hat{\psi}=E_{\hat{\pi}_{t-1}}(\psi)$ and $\Sigma=Var_{\hat{\pi}_{t-1}}(\psi)$. For $i=1,...,N$ set
\begin{align*}
m(\psi^{(i)}) & =a\psi^{(i)}+(1-a)\overline{\psi}\\
v(\psi^{(i)}) & =(1-a^2)\Sigma
\end{align*}
and
\begin{align*}
\alpha(\psi^{(i)}) & =\frac{m(\psi^{(i)})^2}{v(\psi^{(i)})} \\
\beta(\psi^{(i)}) & =\frac{m(\psi^{(i)})}{v(\psi^{(i)})}
\end{align*}
For $k=1,...,N$
\begin{itemize}
\item Draw $I_{k}$ with $P(I_{k}=i) \propto w_{t-1^{(i)}}f_{N}(y_{t}|g(x_{t}^{(i)}),m(\psi^{(i)}))$ where for simplicity $g(x_{t}^{(i)})=E(x_{t}|x_{t-1},m(\psi^{(i)}))$
\item Draw $\psi^{(k)} \sim G(\alpha(\psi^{(I_{k})}),\beta(\psi^{(I_{k})}))$
\item Draw $x_{t}^{(k)} \sim N(x_{t-1}^{(I_{k})},\tau^{{2}^{(k)}})$
\item Set $\tilde{w}_{t}^{k}=\frac{f_N(y_{t}|x_{t}^{(k)},\psi=\psi^{(k)})}{f_N(y_{t}|g(x_{t}^{(I_{k})}),\psi=m(\psi)^{(I_k)})}$
\end{itemize}
\item Normalize the weights
\item Compute the effective sample size ($ESS$)
\item If $ESS<N/2$, resample:
\begin{itemize}
\item Draw a sample of size N, $x_{t}^{(1)},...,x_{t}^{(N)}$,from the discrete distribution $P((x_{t},\psi)=(x_{t}^{(i)},\psi^{(i)}))=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{itemize}
